---
title: "Calculate lagged correlation between two time-series data"
author:
- name: Xiaotao Shen (https://www.shenxt.info/)
date: "Created on 2021-12-25 and updated on `r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: no
  pdf_document:
    toc: no
vignette: >
  %\VignetteIndexEntry{lagged_correlation}
  %\VignettePackage{massqc}
  % \VignetteEngine{knitr::rmarkdown}
  % \usepackage[utf8]{inputenc}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE, echo=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = TRUE,
  out.width = "100%"
)
```

## **Data preparation**

`laggedcor` needs two time-series data and their corresponding time.

* x: time-series 1. numeric vector. 

* y: time-series 2. numeric vector. 

* time1: time for x. 

* time2: time for y.

The demo data can be from `laggedcor`.

```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
library(laggedcor)
data("step_data", package = "laggedcor")
data("heart_data", package = "laggedcor")

head(step_data)
head(heart_data)
```
```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
time_plot(x = step_data$step, time = step_data$time)
```

```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
time_plot(x = heart_data$heart, time = heart_data$time)
```

## **Calculate lagged correlation**

Then we calculate the lagged correlations between step and heart rate.

```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
dim(step_data)
dim(heart_data)
```
The time-series data with less number of time point should be set as `x`.

```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
x = step_data$step
time1 = step_data$time

y = heart_data$heart
time2 = heart_data$time
```


```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
result = 
  calculate_lagged_correlation(x = x, 
                               y = y,
                               time1 = time1, 
                               time2 = time2, 
                               time_tol = 0.2, 
                               step = 2 / 60, 
                               min_matched_sample = 10, 
                               threads = 5, 
                               cor_method = "spearman")
```

The parameters of `calculate_lagged_correlation`;

* `x`: time-series data 1.

* `y`: time-series data 2.

* `time1`: time-series data time 1.

* `time2`: time-series data time 2.

* `time_tol`: time tolerance for time shift. Unit is hour. 0.2 means ± 0.2 hour.

* `step`: step means the matching window, unit is hour.

* `min_matched_sample`: Minimum for match sample number.

* `threads`: thread number.

* `cor_method`: spearman or pearson.


## **Result of lagged correlation**

```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
result
```

We can see that the index of max correlation is 11 and index of global correlation is 12, means that the max correlation is achieved from a shifted time.

## **Result of lagged correlation**

```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
extract_max_cor(object = result)
extract_global_cor(object = result)
```

We can see that when the shift time is "(-1.5,-0.5]", we get the max correaltion: 0.58. This means x (step) changes before y (heart rate).

```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
extract_all_cor(result)
```

```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
extract_all_cor_p(result)
```

```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
extract_shift_time(result, numeric = TRUE)
extract_shift_time(result, numeric = FALSE)
```

## **Evaluate the quality of lagged correlation**

```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
evaluate_lagged_cor(object = result, plot = TRUE)
```

## **Some results**

```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
lagged_scatter_plot(object = result, hex = TRUE, which = "global")
lagged_scatter_plot(object = result, hex = TRUE, which = "max")
```

```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
lagged_alignment_plot(object = result, which = "global", x_limit = c(1,50000))
```